---
output: pdf_document
---
# Reproducible Research: Peer Assessment 1

## Code for reading in the dataset and/or processing the data
```{r loaddata}
unzip(zipfile="activity.zip")
data <- read.csv("activity.csv")
```

## Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, xlab="Mean total number of steps taken each day", binwidth=1000)
# Mean number of steps taken each day
mean(total.steps, na.rm=TRUE)
# Median and median number of steps taken each day
median(total.steps, na.rm=TRUE)
```

## Time series plot of the average number of steps taken
```{r}
library(ggplot2)
av.steps <- tapply(data$steps, data$date, FUN=mean,  na.rm=TRUE)
q<-data.frame(unique(data$date),unname(av.steps))
colnames(q) <- c("Date",'Steps')
q <- na.omit(q)
q$Date <- as.Date(q$Date)
ggplot(data=q,aes(x=Date,y=Steps))+geom_line()+ylab("average number of steps taken")

averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
```

The 5-minute interval that, on average, contains the maximum number of steps
```{r}
averages[which.max(averages$steps),]
```

## Code to describe and show a strategy for imputing missing data


```{r how_many_missing}
missing <- is.na(data$steps)
# How many missing
table(missing)

# Replace each missing value with the mean value of its 5-minute interval
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```
Histogram of the total number of steps taken each day after missing values are imputed

```{r}
total.steps = tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=weekday.or.weekend)
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
```